Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-26 paper/paper.tex
normal continuation -> continuation

We apply this idea to AAM and our first step of transformations is to linearize all
nondeterministic choices. This step removes all nondeterminism from the @step@ function.
Likewise, the @Frame@ saves the information of the caller in concrete executions.
We then introduce another layer of control, and define a case class @NDCont@
that saves the information at a fork point when we have multiple target closures.
We also add a new field @ndk@ to the definition of state, which we now call @NDState@.
For clarity of presentation in differentiating between the two continuations, we elect to call
the first a \emph{normal continuation}, and the second a \emph{meta-continuation}
\footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
continuation} \cite{Danvy:1990:AC:91556.91622}.}.

@@

If the normal continuation @konts@ is an empty list, then it means we have reached
the end of the computation (i.e., a halt); otherwise, we should return the values of 
@ae@ to its caller which is contained in the top frame of stack.

@@

The @drive@ function is also changed: there is no worklist anymore, as all the
potentially unexplored states are embedded into the continuation for
nondeterminism.
The termination of the analysis occurs when we reach an @NDState@ in which the expression
is atomic and both the normal continuation and nondeterminism continuation are
empty lists. This corresponds to the case that @todo@ list is empty.

@@

Function @continue@ is mainly what we have seen for the second case in @drive_step@,
except that @ndcontinue@ is called when the normal continuation is empty.
It works on the list representation of context: An empty list @Nil@ represents the
halt of a computation path, whereas a cons @::@ means the top frame of the list
holds the binding variable and context.
An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.

@@

For clarity, we first show how to refunctionalize normal continuations and
meta-continuation without collecting semantics.
Then, adding an appropriate caching algorithm to make sure the analysis will
terminate at some fixed-point is a simple transformation of the program to cache-passing
style. Finally, we examine how computable pushdown control-flow analysis is established
through refunctionalization and caching.

@@

Readers familiar with functional programming may think of @fold@ at this point; indeed,
it is a variant implementation of @fold@ in continuation-passing style. But here we have two
continuations, one for the next element in the collection, and one for the eventual result.
This is sometimes referred to as an extended continuation-passing style (ECPS)
that function @k@ is a normal continuation and @m@ is a meta-continuation 
\cite{Danvy:1990:AC:91556.91622}. \\

>>>>

We apply this idea to AAM and our first step of transformations is to linearize all
nondeterministic choices. This step removes all nondeterminism from the @step@ function.
Likewise, the @Frame@ saves the information of the caller in concrete executions.
We then introduce another layer of control, and define a case class @NDCont@
that saves the information at a fork point when we have multiple target closures.
We also add a new field @ndk@ to the definition of state, which we now call @NDState@.
For clarity of presentation in differentiating between the two continuations, we elect to call
the first a \emph{continuation}, and the second a \emph{meta-continuation}
\footnote{Traditionally, they are also called \emph{success continuation} and \emph{failure
continuation} \cite{Danvy:1990:AC:91556.91622}.}.

@@

If the continuation @konts@ is an empty list, then it means we have reached
the end of the computation (i.e., a halt); otherwise, we should return the values of 
@ae@ to its caller which is contained in the top frame of stack.

@@

The @drive@ function is also changed: there is no worklist anymore, as all the
potentially unexplored states are embedded into the continuation for
nondeterminism.
The termination of the analysis occurs when we reach an @NDState@ in which the expression
is atomic and both the continuation and meta-continuation are empty lists. This corresponds 
to the case that @todo@ list is empty.

@@

Function @continue@ is mainly what we have seen for the second case in @drive_step@,
except that @ndcontinue@ is called when the continuation @konts@ is empty.
It works on the list representation of context: An empty list @Nil@ represents the
halt of a computation path, whereas a cons @::@ means the top frame of the list
holds the binding variable and context.
An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.

@@

For clarity, we first show how to refunctionalize continuations and
meta-continuation without collecting semantics.
Then, adding an appropriate caching algorithm to make sure the analysis will
terminate at some fixed-point is a simple transformation of the program to cache-passing
style. Finally, we examine how computable pushdown control-flow analysis is established
through refunctionalization and caching.

@@

Readers familiar with functional programming may think of @fold@ at this point; indeed,
it is a variant implementation of @fold@ in continuation-passing style. But here we have two
continuations, one for the next element in the collection, and one for the eventual result.
This is sometimes referred to as an extended continuation-passing style (ECPS)
that function @k@ is a continuation and @m@ is a meta-continuation 
\cite{Danvy:1990:AC:91556.91622}. \\
